Physcraper: A Python package for continually updated gene trees

### 2020-05-04 physcraper_ms.Rmd
summary

Update trees that were reported on a published peer-reviewed study, uploaded to Treebase and to the Open Tree of Life database

@@

@@

Concerns:
Errors in identification of sequences
Little control along the process
Too much of a black box?

@@

The goal of Physcraper is to build upon previous phylogenegtic knowledge,
allowing a direct comparison of existing phylogenies to phylogenies constructed
using new information obtained from GenBank

To achieve this, Physcraper uses the Open Tree of Life phylesystem and connects it
to the TreeBase database, to use the original DNA dataset matrices that produced
a phylogeny in the OToL database as
a starting point to get new genetic data belonging to the focal group of study, and
update the phylogenetic relationships displayed by that dataset in particular.

Physcraper also allows node by node comparison of the trees, using the conflict API of OToL.

@@

The original tree is used as starting tree for the phylogenetic searches, but it
can also be set as a full topological constraint or not used at all, depending on
the goals of the user.

>>>>

1. Phylogenies are still a key part of research in all areas of biology. The nonspecialist
can use one of the many tools available to automatize phylogenetic reconstruction.
However, interpretation of results, comparison with previously available phylogenetic
hypotheses, and choosing of one phylogeny for discussion still impose difficulties
to one that is not a specialist on phylogenetic methods or on a particular group of study.

1. Physcraper is an open‐source Python program that automatizes the update of public
phylogenies by making use of public DNA sequence data and taxonomic information.

1. Physcraper can be used by the nonspecialist, as a tool to generate phylogenetic
hypothesis based on public expert phylogentic knowledge.
The phylogenetics and group specialist will find it useful as a tool to facilitate comparison
of alternative phylogenetic hypotheses (topologies).
***Is physcraper intended for the nonspecialist?? We have two types of nonspecialists:
the ones that do not know about phylogenetic methods and the ones that might know
about phylogenetic methods but do not know of a certain biological group.***

1. Physcraper implements node by node comparison of the the original and the updated
trees using the conflict API of OToL.

1. We hope the physcraper workflow encourages researchers on their data sharing practices.

1. Physcraper can be used with any OS. Detailed instructions for installation and
use are available at <https://github.com/McTavishLab/physcraper>.

@@

There is a lot of phylogenetic knowledge already available.
Some of it has been made public at OToL.
Most or all public phylogenies have the potential to be updated with the ever increasing
amount of genetic data that is available on GenBank.
There are now various methods to automatize use of GenBank data for phylogenetic reconstruction, which are
generally intended to facilitate the use of phylogenetic knowledge for the nonspecialist.
However, none of them provide a means to compare with previously available expert
knowledge on a particular group.

A key aspect of the standard phylogenetic workflow is comparison with already existing
phylogenetic hypotheses and with phylogenies that are considered "best" by experts
not only in phylogenetics, but also experts of the focal group of study.

@@

Concerns with these tools:
Errors in identification of sequences
Little control along the process
Too much of a black box?

@@

The goal of Physcraper is to build upon previous phylogenetic knowledge,
allowing a direct comparison of existing phylogenies to phylogenies that are constructed
using new genetic data from GenBank

To achieve this, Physcraper uses the Open Tree of Life phylesystem and connects it
to the TreeBase database, to (1) get the original DNA dataset matrices (alignments) that produced
a phylogeny that was published and then made available in the OToL database, (2)
use this DNA alignments as a starting point to get new genetic data belonging
to the focal group of study, to (3) finally update the phylogenetic relationships in the group.

@@

The original tree is by defualt used as starting tree for the phylogenetic searches, but it
can also be set as a full topological constraint or not used at all, depending on
the goals of the user.

Physcraper implements node by node comparison of the the original and the updated trees, using the conflict API of OToL.

# Physcraper
